### 2018-02-13 manuscript/02_methods.Rmd
Revise methods
- Add info on flux trait temperature corrections
- Add info on CLM values
- Replace repository link with Open Science Framework link
- Minor text revisions

Foliar trait data for this analysis comes from the TRY global traits database (@kattge_try_2011; Tab.\ \@ref(tab:trydata)).
We focused our research on seven foliar traits:
Leaf longevity (months),
specific leaf area (SLA, m$^2$ kg$^{-1}$),
leaf nitrogen content ($N_{mass}$, mg N g$^{-1}$ or $N_{area}$, g m$^{-2}$),
leaf phosphorus content ($P_{mass}$, mg P g$^{-1}$ or $P_{area}$, g m$^{-2}$),
leaf dark respiration at leaf temperature ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
maximum Rubisco carboxylation rate at 25°C ($V_{c,max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $V_{c,max,area}$, µmol m$^{-2}$ s$^{-1}$),
and maximum electron transport rate ($J_{max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{max,area}$, µmol m$^{-2}$ s$^{-1}$).
We performed analyses separately for both mass- and area-normalized traits to avoid potential artifacts caused by different trait normalization [@osnas_global_2013; @lloyd_les].
We restricted our analysis to TRY data that have been quality-controlled and for which adequate species information was provided for functional type classification [see @kattge_try_2011].

@@

We assigned each species to a unique plant functional type (PFT) following the scheme in the Community Land Model [CLM4.5, @clm45_note; Tab.\ \@ref(tab:pfts), Fig.\ \@ref(fig:samplesize)].
We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from the TRY database.
Where species attributes disagreed between datasets (e.g., categorized as a shrub in one dataset but a tree in another), we assigned the attribute that was observed most frequently between the datasets (e.g., if five datasets say "shrub" but only one says "tree", we would classify it as a shrub).
Where species lacked data on certain attributes, we assigned those attributes based on higher order phylogeny where appropriate (e.g., \em{Poaceae} family are usually grasses, \em{Larix spp.} are deciduous needleleaved trees) or otherwise omitted the species from our analyses.
For biome specification, we collected all latitude and longitude data for each species,
matched these data to 30 second ($\sim1$ km$^{2}$) mean annual temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 [@worldclim],
calculated the mean AMT for all sites at which that species was observed,
and then binned these species based on the following cutoffs: boreal/arctic ($AMT \leq 5^\circ C$), temperate ($AMT \leq 20^\circ C$), and tropical ($AMT > 20^\circ C$).

@@

For each model fit, we ran five parallel MCMC chains, automatically increasing the sample size until the final result achieved convergence as determined by a Gelman-Rubin potential scale reduction statistic less than 1.1 [@gelman_1992_inference].
We implemented this sampling algorithm in an open source, publicly available R [version 3.4.2; @rstats] package (<http://github.com/ashiklom/mvtraits>).

@@

To assess the impact of multivariate and hierarchical constraint on trait estimates,
we compared the mean and 95% confidence intervals of trait estimates for each PFT from each model (Fig.\ \@ref(fig:meancompare)).

@@

We then fit a generalized linear model relating relative uncertainty to sample size ($n$) for each of the model types (univariate, multivariate, and hierarchical; Fig. \@ref(fig:civssamplesize)).

@@

To assess the consistency of among- and within-PFT trait trade-offs, we looked at covariance estimates for each trait pair and, where these values were significantly different from zero ($p < 0.05$),
we calculated the eigenvalues from the pairwise variance-covariance matrix for that trait pair and plotted the corresponding dominant eigenvectors centered on the mean estimates (Fig.\ \@ref(fig:stickpairs)).
This figure provides a visual representation of relative positions of PFTs in trait space and both the direction and extent of within-PFT trait covariance, and is directly analogous to conceptual figures describing hierarchical trait variability across environmental gradients as presented in, for instance, @cornwell_community_2009 and @albert_intraspecific_2010.
Due to the small number of points used to estimate across-PFT covariance in the hierarchical model, none of the across-PFT covariances estimated in the hierarchical model were significantly different from zero ($p < 0.05$).
For this reason, we compared within-PFT covariances as estimated by the hierarchical model with the covariances estimated by fitting a multivariate model to all of the data.

Besides the consistency in the direction of trait covariance globally and between different PFTs, we also investigated the strength and predictive power of these covariances, which is represented by correlation coefficients (i.e. the pairwise covariance normalized to the variances of the component variables).
To do this, we plotted the mean and 95% confidence interval of the pairwise trait correlation coefficients for the global estimate from the pooled multivariate model and PFT-level estimates from the hierarchical model (Fig.\ \@ref(fig:corrbars)).

The R code for running these analyses is publicly available online at <http://github.com/ashiklom/np_trait_analysis>.
The TRY data used for this analysis can be requested at <http://try-db.org>.

>>>>

Foliar trait data for this analysis comes from the TRY global traits database (@kattge_try_2011; Tab.\ \@ref(tab:trydata)).
We focused our research on seven foliar traits:
Leaf longevity (months),
specific leaf area (SLA, m$^2$ kg$^{-1}$),
leaf nitrogen content ($N_{mass}$, mg N g$^{-1}$ or $N_{area}$, g m$^{-2}$),
leaf phosphorus content ($P_{mass}$, mg P g$^{-1}$ or $P_{area}$, g m$^{-2}$),
leaf dark respiration at 25°C ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
maximum Rubisco carboxylation rate at 25°C ($V_{c,max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $V_{c,max,area}$, µmol m$^{-2}$ s$^{-1}$),
and maximum electron transport rate at 25°C ($J_{max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{max,area}$, µmol m$^{-2}$ s$^{-1}$).
For $V_{c,max}$, we only used values already reported in TRY as being at 25°C.
For $R_{d}$, we normalized the values to 25°C based on reported leaf temperature values following the same methods as @atkin_global_2015.
For $J_{max}$, we normalized the values to 25°C based on reported leaf temperature values using the temperature response function described in @kattge_2007_temperature (Equation 1 therein).
To avoid potential artifacts caused by different trait normalization, we performed analyses separately for both mass- and area-normalized traits [@osnas_global_2013; @lloyd_les].
We restricted our analysis to TRY data that have been quality-controlled and for which adequate species information was provided for functional type classification [see @kattge_try_2011].

@@

We assigned each species to a unique plant functional type (PFT) following the scheme in the Community Land Model [CLM4.5, @clm45_note; Tab.\ \@ref(tab:pfts), Fig.\ 1].
We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from the TRY database.
Where species attributes disagreed between datasets (e.g., categorized as a shrub in one dataset but a tree in another), we assigned the attribute that was observed most frequently between the datasets (e.g., if five datasets say "shrub" but only one says "tree", we would classify it as a shrub).
Where species lacked data on certain attributes, we assigned those attributes based on higher order phylogeny where appropriate (e.g., \em{Poaceae} family are usually grasses, \em{Larix spp.} are deciduous needleleaved trees) or otherwise omitted the species from our analyses.
For biome specification, we collected all latitude and longitude data for each species,
matched these data to 30 second ($\sim1$ km$^{2}$) mean annual temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 [@worldclim],
calculated the mean AMT for all sites at which that species was observed,
and then binned these species based on the following cutoffs: boreal/arctic ($AMT \leq 5^\circ C$), temperate ($AMT \leq 20^\circ C$), and tropical ($AMT > 20^\circ C$).

@@

For each model fit, we ran five parallel MCMC chains, continuing the sampling until the final result achieved convergence as determined by a Gelman-Rubin potential scale reduction statistic less than 1.1 [@gelman_1992_inference].
We implemented this sampling algorithm in an open source, publicly available R [version 3.4.3; @rstats] package (<http://github.com/ashiklom/mvtraits>).

@@

To assess the impact of multivariate and hierarchical constraint on trait estimates,
we compared the mean and 95% confidence intervals of trait estimates for each PFT from each model (Fig.\ 2).
For reference, we also added the default parameter values of CLM 4.5 [Table 8.1 in @clm45_note] for SLA, $N_{mass}$, $N_{area}$, $V_{c,max,mass}$, and $V_{c,max,area}$ to Fig.\ 2.
To convert CLM's reported C:N ratio to $N_{mass}$, we assumed a uniform leaf C fraction of 0.46.
We then divided this calculated $N_{mass}$ by the reported SLA to obtain $N_{area}$.
We calculated $V_{c,max,mass}$ by multiplying the reported $V_{c,max,area}$ by the reported SLA.

@@

We then fit a generalized linear model relating relative uncertainty to sample size ($n$) for each of the model types (univariate, multivariate, and hierarchical; Fig. 3).

@@

To assess the consistency of among- and within-PFT trait trade-offs, we looked at covariance estimates for each trait pair and, where these values were significantly different from zero ($p < 0.05$),
we calculated the eigenvalues from the pairwise variance-covariance matrix for that trait pair and plotted the corresponding dominant eigenvectors centered on the mean estimates (Fig.\ 4).
This figure provides a visual representation of relative positions of PFTs in trait space and both the direction and extent of within-PFT trait covariance, and is directly analogous to conceptual figures describing hierarchical trait variability across environmental gradients as presented in, for instance, @cornwell_community_2009 and @albert_intraspecific_2010.
Due to the small number of points used to estimate across-PFT covariance in the hierarchical model, none of the across-PFT covariances estimated in the hierarchical model were significantly different from zero ($p < 0.05$).
For this reason, we compared within-PFT covariances as estimated by the hierarchical model with the covariances estimated by fitting a multivariate model to all of the data.

Besides the consistency in the direction of trait covariance globally and between different PFTs, we also investigated the strength and predictive power of these covariances, which is represented by correlation coefficients (i.e. the pairwise covariance normalized to the variances of the component variables).
To do this, we plotted the mean and 95% confidence interval of the pairwise trait correlation coefficients for the global estimate from the pooled multivariate model and PFT-level estimates from the hierarchical model (Fig.\ 5).

The R code and ancillary data for running these analyses is publicly available online via the Open Science Framework (OSF) at <https://osf.io/w8y73/>.
The TRY data used for this analysis can be requested at <http://try-db.org>.
